---
title: "NYPD Shooting Incident Data Analysis"
date: "`r Sys.Date()`"
output: html_document
---




install.packages("ggplot2")
install.packages("janitor")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(tsibble)
library(cluster)
library(factoextra)
```

## 1. Data load
This data get from here "https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic"
```{r}
data <- read_csv("NYPD_Shooting_Incident_Data__Historic_.csv")
glimpse(data)
summary(data)
```

## 2. Trends in the number of incidents by year

```{r}
data$OCCUR_DATE <- mdy(data$OCCUR_DATE)
data$YEAR <- year(data$OCCUR_DATE)

yearly_data <- data %>% 
  count(YEAR)

ggplot(yearly_data, aes(x = YEAR, y = n)) +
  geom_line() +
  geom_point() +
  labs(title = "Yearly Shooting case count", x = "Year", y = "Case")
```

## 3. Monthly and weekly trends

```{r}
data$MONTH <- month(data$OCCUR_DATE, label = TRUE)
data$WEEKDAY <- wday(data$OCCUR_DATE, label = TRUE)

ggplot(data, aes(x = MONTH)) +
  geom_bar() +
  labs(title = "Monthly Shooting case count", x = "Month", y = "Case")

ggplot(data, aes(x = WEEKDAY)) +
  geom_bar() +
  labs(title = "Weekly Shooting case count", x = "Week", y = "Case")
```

- Although between 2010-2019 declined, 2020 sharply increased. Recently, it has declined again.
- Furthermore, incidents tend to increase between May and September, with most occurring 1.5 times on weekends.


## 4. Number of incidents by district (BORO)

```{r}
ggplot(data, aes(x = BORO)) +
  geom_bar(fill = "steelblue") +
  labs(title = "BORO Shooting case count", x = "BORO", y = "case") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
- Concentrated in specific locations, there seem to be hotspots.


## 5. Analysis about Victime&Perpetrator

```{r}
valid_age_groups <- c("<18", "18-24", "25-44", "45-64", "65+", "UNKNOWN")

data %>%
  filter(
    !is.na(VIC_AGE_GROUP),
    VIC_AGE_GROUP %in% valid_age_groups  # ← この行が重要
  ) %>%
  mutate(VIC_AGE_GROUP = factor(VIC_AGE_GROUP, levels = valid_age_groups)) %>%
  ggplot(aes(x = VIC_AGE_GROUP)) +
  geom_bar(fill = "gray30") +
  labs(title = "Victim age", x = "Age", y = "Case") +
  theme_minimal()

ggplot(data, aes(x = VIC_SEX, fill = VIC_RACE)) +
  geom_bar(position = "dodge") +
  labs(title = "Victim race&sex", x = "Sex", y = "Case")

valid_perp_age_groups <- c("<18", "18-24", "25-44", "45-64", "65+", "UNKNOWN")

data %>%
  filter(
    !is.na(PERP_AGE_GROUP),
    PERP_AGE_GROUP %in% valid_perp_age_groups
  ) %>%
  mutate(PERP_AGE_GROUP = factor(PERP_AGE_GROUP, levels = valid_perp_age_groups)) %>%
  ggplot(aes(x = PERP_AGE_GROUP)) +
  geom_bar(fill = "gray30") +
  labs(title = "Perpetrator age", x = "Age", y = "Case") +
  theme_minimal()

data %>%
  filter(
    !is.na(PERP_SEX),
    !is.na(PERP_RACE),
    PERP_SEX != "(null)",
    PERP_RACE != "(null)"
  ) %>%
  ggplot(aes(x = PERP_SEX, fill = PERP_RACE)) +
  geom_bar(position = "dodge") +
  labs(title = "Perpetrator race & sex", x = "Sex", y = "Case") +
  theme_minimal()
```

- Most of the victims were between the ages of 18 and 44, with the majority being black, followed by Hispanic.
- The perpetrators are of the same age group as the victims, but I think there are many cases where the perpetrators are not identified.

## 6. Clustering analysis (K-means + PCA)

```{r}
library(dplyr)
library(factoextra)
library(ggpubr)

set.seed(123)
# Data preprocessing and clustering
clust_data <- data %>%
  select(VIC_SEX, VIC_AGE_GROUP, VIC_RACE, PERP_SEX, PERP_AGE_GROUP, PERP_RACE) %>%
  filter(across(everything(), ~ !is.na(.))) %>%
  mutate(across(everything(), as.factor)) %>%
  slice_sample(n = 1000)

# One-hot encode + constant sequence removal
clust_dummy <- model.matrix(~ . - 1, data = clust_data)
clust_dummy_filtered <- clust_dummy[, apply(clust_dummy, 2, function(x) length(unique(x)) > 1)]

# PCA + K-means
pca_result <- prcomp(clust_dummy_filtered, 
                     center = TRUE, 
                     scale. = TRUE, 
                     rank. = 2)  # PC1〜2

kmeans_result <- kmeans(pca_result$x[, 1:2], centers = 3)

# Plot
fviz_cluster(
  list(data = pca_result$x[, 1:2], cluster = kmeans_result$cluster),
  geom = "point",
  main = "K-means Clustering（+ PCA）"
)

# Estimating the optimal number of clusters using the elbow method
fviz_nbclust(pca_result$x[, 1:2], kmeans, method = "wss") +
  labs(title = "the elbow method")

```

- After performing PCA (Component=2), we confirmed the trend using k-mean clustering.
The elbow method showed that three clusters were most appropriate, and we were able to successfully divide the data into three clusters.


## 8. Clustering analysis (K-means + PCA)
- In order to confirm which categories each cluster corresponds to, we checked the weights of each feature in each axis (PC1, 2) and cluster (n=1-3).
```{r}
# 1. Combining one-hot encoded data with clusters
clust_df <- as.data.frame(clust_dummy_filtered)
clust_df$cluster <- factor(kmeans_result$cluster)

# 2. Calculate the average (ratio) for each cluster.
cluster_means <- clust_df %>%
  group_by(cluster) %>%
  summarise(across(everything(), mean)) %>%
  ungroup()

# 3. Convert to tidy format
cluster_long <- cluster_means %>%
  pivot_longer(cols = -cluster, names_to = "Feature", values_to = "Proportion")

# 4. Heatmap
ggplot(cluster_long, aes(x = cluster, y = fct_rev(Feature), fill = Proportion)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", Proportion)), size = 3) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(
    title = "Cluster × feature ratio",
    x = "cluster", y = "feature", fill = "ratio"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(face = "bold"),
        axis.text.y = element_text(size = 8))


# 1. PCA loading
loadings <- as.data.frame(pca_result$rotation)

# 2. Convert to tidy format
loadings_long <- loadings %>%
  mutate(Feature = rownames(.)) %>%
  pivot_longer(cols = starts_with("PC"), names_to = "Component", values_to = "Loading")

# 3. Heatmap
ggplot(loadings_long, aes(x = Component, y = fct_rev(Feature), fill = Loading)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", Loading)), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "The loading in each principal component",
    x = "PC", y = "feature", fill = "Loading"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(face = "bold"),
        axis.text.y = element_text(size = 8))

```

- PC1 is the axis that mainly contains information about perpetrators, while PC2 shows a combination of information about victims and perpetrators.
- ①Cluster 1 consists of cases where the victim is black and the perpetrator is also black.
- ②Cluster 2 consists of cases where the victim is black&Hispanic and the perpetrator is Black.
- ③Cluster 3 consists of cases where the victim&perpetrator is Hispanic.

## Conlusion

- There is a clear trend in the months and days of the week when shooting incidents occur most frequently.
- There are also hot spots where they occur.
- Clustering reveals patterns in the attributes of victims and perpetrators.As a point for future consideration, I believe it should examine the racial composition of the population in the incident areas and compare it with other regions. 



